---
title: "bioRxiv preprint stats"
author:
- "Hao Ye"
- "Michael Parkin"
date: "`r Sys.Date()`"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache = TRUE)
```

# Setup

```{r}
library(tidyverse)
```

# Data Overview

See the `compute_paper_stats.R` file for the code used to generate the paper stats data.

The data collects information on preprints, with one row per preprint. Stats are computed by summing up monthly views and downloads for up to the first 3 months after posting date (posted_date + 3 months >= traffic_date >= posted_date). Traffic stats are collected monthly, so we assign a date equal to the last day of the month for the traffic.

Columns are:

* `id` - unique numerical id for the preprint
* `downloads` - number of pdf downloads
* `views` - number of views of the abstract
* `traffic_date` - latest traffic date for the summed views and downloads
* `posted_date` - date of initial posting to bioRxiv
* `collection` - the subject area designation on bioRxiv
* `duration` - the # of days, computed as traffic_date - posted_date + 1
* `downloads_per_day` - downloads / duration
* `views_per_day` - views / duration
* `pub_date` - date of publication of preprint (NA if not yet published)

```{r}
paper_stats <- readRDS("paper_stats.RDS")
str(paper_stats)
```

## Filter papers

We want to focus on "low-visibility" preprints, so use only the papers that are unpublished or published after the last date of traffic collection (posted date + 3 months).

```{r}
paper_stats <- paper_stats %>%
    mutate(unpublished = is.na(pub_date) | 
               pub_date > traffic_date)
```

## Examine statistical quantiles of views per day

Compute quantiles:
```{r}
quantile_vec <- c(0.05, 0.1, 0.25, 0.5, 0.75, 0.9, 0.95)

views_quantiles <- paper_stats %>%
    filter(unpublished) %>%
    select(collection, views_per_day, downloads_per_day) %>%
    nest(-collection) %>%
    mutate(views_q = map(data, ~ quantile(.$views_per_day, probs = quantile_vec)), 
           views_q = map(views_q, ~ bind_rows(.) %>% gather())) %>%
    select(-data) %>% 
    unnest() %>%
    mutate(key = factor(key, levels = c("5%", "10%", "25%", "50%", "75%", "90%", "95%"))) %>%
    spread(key, value)

knitr::kable(views_quantiles)
```

## Distributions of views per day

```{r, fig.width = 6.5, fig.height = 18}
paper_stats %>%
    ggplot(aes(x = views_per_day, color = unpublished)) + 
    geom_density() + 
    coord_cartesian(xlim = c(0, 100)) + 
    facet_wrap(~collection, scales = "free", ncol = 4) + 
    theme_bw(base_size = 12) + 
    theme(axis.text.x = element_text(angle = 90, hjust = 1), 
          legend.position  = "top")
```

## How does recency of a preprint affect views

With more preprints being posted, we might expect the # of views per day to change with recency. For the least-visible preprints, the number of views per day seems to be mostly constant. However, depending on category, the upper-quantiles can vary, sometimes decreasing with recency if sample sizes are small.

```{r, fig.width = 6.5, fig.height = 18}
paper_stats %>%
    ggplot(aes(x = posted_date, y = views_per_day)) + 
    geom_point() + 
    geom_quantile(quantiles = quantile_vec, formula = "y ~ x") + 
    facet_wrap(~collection, scales = "free", ncol = 4) + 
    theme_bw(base_size = 12)
```

## How does traffic decrease over time for a sample of preprints

```{r}
db <- readRDS("raw_stats.RDS")

set.seed(42)
selected_ids <- sample(db$id, 50)

db %>%
    filter(id %in% selected_ids) %>%
    mutate(id = factor(id)) %>%
    group_by(id) %>%
    ggplot(aes(x = traffic_date - posted, y = views, color = id)) + 
    geom_line() + 
    theme_bw(base_size = 20) + 
    guides(color = "none")
```

# Compute an "Invisibility" index

Compute index as fraction of preprints in the same collection with at least as many views_per_day.

```{r}
paper_indices <- paper_stats %>%
    group_by(collection) %>%
    mutate(index = rank(views_per_day, ties.method = "min"), 
           index = 1 - index / n()) %>%
    ungroup() %>%
    select(id, collection, views_per_day, index, 
           traffic_date, posted_date, pub_date, 
           views, downloads)
```

## Select a few papers per collection to display

```{r selected papers per collection}
title_lookup <- db %>%
    select(id, title) %>%
    distinct()

selected_papers <- paper_indices %>%
    group_by(collection) %>%
    top_n(10, index) %>%
    arrange(collection, desc(index)) %>%
    left_join(title_lookup, by = "id") %>%
    mutate(pub = ifelse(is.na(pub_date), "N", "Y")) %>%
    select(collection, index, 
           "views/day" = views_per_day, 
           posted = posted_date, pub, title)

knitr::kable(selected_papers, digits = 3)
```
